---
title: "R Notebook"
output: html_notebook
author: "Andrew Payne"
---
Data Preparation Stage

```{r}
# import our previously prepared dataframe from dataset prep.Rmd

df <- read.csv("D:\\Libraries\\OneDrive - Ryerson University\\Project\\data\\df processed 2.csv")

# unnecessary columns have been removed prior to importing
```

# backup
df_copy <- df

```{r}
# remove rows with NA values: these correspond to games too early in the season
# to establish rolling averages of game performance statistics
df <- na.omit(df)

# verify all NA values are removed
sum(is.na(df)) # 0
```

```{r}
# remove outliers:
# - games where a team has 0 days of rest or >4 days of rest
df <- df[df$teamDayOff < 5 & df$opptDayOff < 5, ]
df <- df[df$teamDayOff > 0 & df$opptDayOff > 0, ]

# - games with more or less than 240 minutes played
df <- df[df$teamMin == 240 & df$opptMin == 240, ]

# - games with non-zero values in the team/opptPTS5-8 columns
df <- df[df$teamPTS5 == 0 & df$opptPTS5 == 0, ]
# df[df$teamPTS6 != 0 | df$opptPTS6 != 0, ] returns 0 rows
# df[df$teamPTS7 != 0 | df$opptPTS7 != 0, ] returns 0 rows
# df[df$teamPTS8 != 0 | df$opptPTS8 != 0, ] returns 0 rows

# the teamMin, opptMin, and team/opptPTS5-8 columns can now be discarded

df <- subset(df, select = -c(teamMin, opptMin, teamPTS5, teamPTS6, teamPTS7,
                             teamPTS8, opptPTS5, opptPTS6, opptPTS7, opptPTS8))
```

# backup
df_copy2 <- df

```{r}
# correct a minor issue with %-based factors
# (percentages were computed as rolling averages when they should be computed
#  based on [3P / FT] FGs made / [3P / FT] FGs attempted)

for (game_id in 1:nrow(df)) {
  # FG%
  df$teamFG.[game_id] <- round(df$teamFGM[game_id] / df$teamFGA[game_id], 3) * 100
  df$opptFG.[game_id] <- round(df$opptFGM[game_id] / df$opptFGA[game_id], 3) * 100
  
  # 3FG%
  df$team3P.[game_id] <- round(df$team3PM[game_id] / df$team3PA[game_id], 3) * 100
  df$oppt3P.[game_id] <- round(df$oppt3PM[game_id] / df$oppt3PA[game_id], 3) * 100
  
  # FT%
  df$teamFT.[game_id] <- round(df$teamFTM[game_id] / df$teamFTA[game_id], 3) * 100
  df$opptFT.[game_id] <- round(df$opptFTM[game_id] / df$opptFTA[game_id], 3) * 100
}
```

```{r}
# factorize class variable 'teamRslt'

df$teamRslt <- as.factor(df$teamRslt)
```

```{r}
# normalize the numeric factors of the dataset
normalize <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

df_norm <- as.data.frame(lapply(subset(df, select = -c(teamRslt)), normalize))
df_norm$teamRslt <- df$teamRslt
```

#backup
df_copy3 <- df

```{r}
# feature extraction - identify the most important factors of the dataset
install.packages("caret")
library(caret)

# remove redundant features from model consideration

# calculate correlation matrix for all factors except class variable
correlationMatrix <- cor(df_norm[,1:42])
highCorrelation <- findCorrelation(correlationMatrix, cutoff=0.75, names = TRUE)

# remove highly correlated factors from data frame, based on cor()
df_redundancy <- df_norm[setdiff(names(df_norm), highCorrelation)]
```

set.seed(1)

```{r}
# k-fold cross validation model with Naive Bayes classifier

library(caret)

cross_val <- trainControl(method = "cv",
                          number = 10,)

model <- train(teamRslt ~ ., data = df_redundancy,
               method = "naive_bayes", trControl = cross_val)

# methods = logistic regression ("logreg"), Naive Bayes ("naive_bayes", "nb")

model # 65.75% average accuracy
```